---
title: "Shiny App Demo"
runtime: shiny
---

# Intro to Shiny Widgets

<div class="alert alert-info">
This content was created by Dr. Jesse Lecy as part of the ["Intro to data science"](https://ds4ps.org/course_website/assignments/#labs) class at ASU
</div>


Shiny apps have two parts: a user interface (page layouts and input functions), and a set of rendering functions that translate R graphics into HTML and javascript elements to create interactive web pages.

* `inputPanel` or `sidebarPanel` creates a panel for the widgets
* `input` functions allow users to specify inputs
* `renderPlot` converts to HTML / javascript

When using Shiny, a global object called "input" will be created, and it stores the input values the user defines in slots created through the `inputId` arguments. For example, below we use "animal" for the `inputID`, so the input values are referenced at `input$animal`.


```{r}

selectInput( 
             inputId='animal', 
             label='Select Animal', 
             choices=c("cat","dog","rat"),
             selected="dog"
           )
    
renderPlot({

    plot.new()
    plot.window( xlim=c(0,10), ylim=c(0,10) )
    box()
    
    text( 5, 5, label=input$animal, cex=4 )  # global object "input"
    
    title( main="My Favorite Animal Is A:", line=-2 )

})

```

Add a panel for the widget.

```{r}

inputPanel(
 
     selectInput( inputId='animal.02', 
                  label='Select Animal', 
                  choices=c("cat","dog","rat"),
                  selected="dog"
                ),
     
     selectInput( inputId='color', 
                  label='Select Color', 
                  choices=c("steelblue","darkred","orange"),
                  selected="orange"
                ),

     selectInput( inputId='size', 
                  label='Select Size', 
                  choices=c(1,2,3,4,5,6,7,8,9),
                  selected=4
                )  
     
) # end inputPanel


renderPlot({

    plot.new()
    plot.window( xlim=c(0,10), ylim=c(0,10) )
    box()
    
    # global object "input" stores all values in slots using their inputId's above
    # inputId='animal.02'
    # inputId='color'
    # inputId='size'
    text( 5, 5, label=input$animal.02, cex=as.numeric(input$size), col=input$color )  
    
    title( main="My Favorite Animal Is A:", line=-2 )

})

```
  
Use the argument `echo=F` in your chunk header to hide the code when you render the widget in a markdown document.  
```{r, echo=F}


 
selectInput( 
             inputId='animal2', 
             label='Select Animal', 
             choices=c("cat","dog","rat"),
             selected="dog"
           )
    




renderPlot({

    plot.new()
    plot.window( xlim=c(0,10), ylim=c(0,10) )
    box()
    
    text( 5, 5, label=input$animal2, cex=4 )
    
    title( main="My Favorite Animal Is A:", line=-2 )

})

```
     

 
 
 
 